Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with autodiff

### 2022-11-23 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/optics.pdf}
    \caption{Summary of the optical configuration. Left panel: The top plot show the \ac{opd} of the pupil at the aperture of the telescope. The binary values create a half-wave step at the mean observation wavelength. The bottom plot shows the resulting large single-star \ac{psf} without aberrations applied. Middle panel: The top plot shows the total \ac{opd} of the pupil with the optical aberrations applied using low-order Zernike polynomials that is used to generate the data. The bottom plot shows the resulting single-star \ac{psf} with the aberrations applied. Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration. Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane. These values have a large spread and would greatly affect any results without calibration. The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
    \label{fig:optics}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/data_resid.pdf}
    \caption{Left: single image of the full four that is used to create the full data set. This image has had both photon and detector noise applied. By eye it is clearly difficult to disentangle the astrophysical information. The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible. Middle: the residual of this single image from the data and the initial uncalibrated model. Clearly these residuals are large, showing that there is a large amount of calibration required. Right: this same residual after the model has been optimised. The residual values are much smaller and at the noise floor. A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
    \label{fig:data_residual}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.}
    \label{fig:astro_params}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/aberrations.pdf}
    \caption{Recovery of the optical aberrations after optimisation. Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals. Middle and right: the true total \ac{opd} and the residuals generated by these values respectively.}
    \label{fig:aberrations}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
    \caption{This figure shows the PRF values after optimisation. The left panel shows the correlation between the pixel-level true and recovered values. Each ploint is color-coded with the total flux incident to each pixel, ie its \ac{snr}. The pixels with a greater signal are recovered better as expected. The right panel shows the histogram of these residuals, with the majority of values being very well recovered and small symmetric about a residual of zero.}
    \label{fig:flat_field}
\end{figure*}

@@

This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}. However, this method does have its limitations. The use of a forwards model clearly requires a good match between the true and model \ac{psf}. The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem. Similarly since we are working with direct image data as opposed to reduced data-products is that miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model.

There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime. Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded. This can be directly seen in Figure~\ref{fig:flat_field}, with the PRF residual having greater variance in for low-flux pixels. This however has a simple solution of observing brighter targets for calibration purposes or observing chosen targets for longer, until a sufficient amount of information is encoded into the data.

It might also seem limited without a dense region of stars to observe as we can not learn about the pixel-level effects in regions with no signal, however we only care about the regions of the detector that have photons landing in them as these are the only ones that will be affected by miss-calibrations in detector level effects.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/optics.pdf}
    \caption{Summary of the optical configuration. Left panel: The top plot show the \ac{opd} of the pupil at the aperture of the telescope. The binary values create a half-wave step at the mean observation wavelength. The bottom plot shows the resulting large single-star \ac{psf} without aberrations applied. Middle panel: The top plot shows the total \ac{opd} of the pupil with the optical aberrations applied using low-order Zernike polynomials that is used to generate the data. The bottom plot shows the resulting single-star \ac{psf} with the aberrations applied. Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration. Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane. These values have a large spread and would greatly affect any results without calibration. The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
    \label{fig:optics}
    \script{gen_optics.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/data_resid.pdf}
    \caption{Left: single image of the full four that is used to create the full data set. This image has had both photon and detector noise applied. By eye it is clearly difficult to disentangle the astrophysical information. The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible. Middle: the residual of this single image from the data and the initial uncalibrated model. Clearly these residuals are large, showing that there is a large amount of calibration required. Right: this same residual after the model has been optimised. The residual values are much smaller and at the noise floor. A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
    \label{fig:data_residual}
    \script{gen_data_resid.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.}
    \label{fig:astro_params}
    \script{gen_astro_params.py}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/aberrations.pdf}
    \caption{Recovery of the optical aberrations after optimisation. Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals. Middle and right: the true total \ac{opd} and the residuals generated by these values respectively.}
    \label{fig:aberrations}
    \script{gen_aberrations.py}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/ff.pdf}
    \caption{This figure shows the PRF values after optimisation. The left panel shows the correlation between the pixel-level true and recovered values. Each ploint is color-coded with the total flux incident to each pixel, ie its \ac{snr}. The pixels with a greater signal are recovered better as expected. The right panel shows the histogram of these residuals, with the majority of values being very well recovered and small symmetric about a residual of zero.}
    \label{fig:flat_field}
    % \script{gen_FF.py}
\end{figure*}

@@

\subsection{Limitations}
This algorithm demonstrates promising performance applied to complex high-dimensional optimisation problems as shown in Section~\ref{sec:phaseretrieval}. However, this method does have its limitations.

% poor model
The use of a forwards model clearly requires a good match between the true and model \ac{psf}. The benefit of using differentiable optical models however is that it provides the perfect platform upon which to directly learn this model from the data, with the ability to scale the complexity as required for each problem. Similarly since we are working with direct image data as opposed to reduced data-products, miss-calibrations of the system are simple to diagnose through direct examination of the residuals of the data and model. Figure~\ref{fig:data_residual} demonstrates that in this example we can be sure of a good fit because the final residual is purely noise

% Single noise realisation
We further examine the limits in which this method begins to break down. When only examining a single image each pixel only has a single realisation of both photon and detector noise, resulting in an over-fitting of the PRF fit to this noise. Furthermore, in the low \ac{snr} regime, the model can fail to converge due to divergent flux values as the model tries to correct for the noise across the whole image. We test this limitation by fitting a single-star using a single image with increases levels of flux from $10^4$\,photons to $10^7$\,photons, using the same optical prescription as described in Section~\ref{sec:simulation}. Figure~\ref{fig:diverge} show the correlation of the recovered fluxes and the fractional flux errors. The low \ac{snr} models diverge from their true values

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/divergence.png}
    \caption{\textcolor{red}{TODO: Caption}}
    \label{fig:diverge}
    % \script{gen_FF.py}
\end{figure*}

In the high \ac{snr} regime and hence the per-pixel freedom recovered With near complete per-pixel freedom and only a single realisation of noise ie when only using a single image, each pixel
Taking the low-flux single-star single image
There are also clear limits in which these methods will begin to fail, such as the low \ac{snr} regime. One

Without the ability to recover the \ac{psf} model accurately, pixel level information will be dominated by the noise processes and the PRF will not be sufficiently encoded. This can be directly seen in Figure~\ref{fig:flat_field}, with the PRF residual having greater variance in for low-flux pixels. This however has a simple solution of observing brighter targets for calibration purposes or observing chosen targets for longer, until a sufficient amount of information is encoded into the data.

\subsection{Other?}
It might also seem limited without a dense region of stars to observe as we can not learn about the pixel-level effects in regions with no signal, however we only care about the regions of the detector that have photons landing in them as these are the only ones that will be affected by miss-calibrations in detector level effects.
